---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

This package provides a set of simple handling functions for importing, parsing, and summarising ERA5 netcdf4 files downloaded from the copernicus climate data store (CDS). It has been tested using single levels and pressure levels data, with files downloaded both manually and through API methods such as [ecmwfr](https://github.com/bluegreen-labs/ecmwfr). 

A simple workflow is outlined below.

1) Create an account on the Copernicus CDS

* Navigate to https://cds.climate.copernicus.eu/cdsapp#!/home and create an account. 

2) Download ERA5 netcdf4 data using manual or API methods

* ERA5 data can be fetched manually from the CDS using the browser interface: https://cds.climate.copernicus.eu/cdsapp#!/search?type=dataset&text=ERA5. 

3) Name files in a local store

* Give the files a suitable name that conforms to your chosen file naming convention. If you are working with individual field types (for example, 2m temperature), but are storing those files alongside other file of different field types, files of a common field type should be named to reflect this.

4) Install and load the ERA5handlers package

* This package and its dependencies can be fetched with the devtools package, using `devtools::install_github("MRPHarris/ERA5handlers")`

5) Collate desired files

* The `collate_era5()` function imports and parses ERA5 netcdf4 files. The package ships with a small netcdf4 file containing 2m temperature data from 1999, used in line with the Copernicus [license](https://cds.climate.copernicus.eu/api/v2/terms/static/licence-to-use-copernicus-products.pdf). The data includes the grid points lat = c(-80.00 -80.25 -80.50) and lon = c(-81.50 -81.25 -81.00 -80.75 -80.50), covering a small portion of the southern Ellsworth Mountains in Antarctica.

In this case, we use `extract_era5` to read in the single file, narrowing the coordinates slightly.

```{r, include = TRUE}
# # Load package
# library(pacman)
# library(ERA5handlers)
# # Specify local data store
# dat_store_era5 <- "D:/DATA/General data/ERA5/"
# # Get all the files in the directory
# era5_fnames <- list.files(dat_store_era5, full.names = TRUE)
# # Get target era5 file
# temp2m_2000_PH <- collate_era5(era5_fnames, string = 'temp2m_1999', coords = c(-80.25, -81.25))
```


